Supplementary Information

### 2020-10-08 writing/transect_supplement.Rmd
add caching in supplement to speed up rendering

datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccccccccicc")
regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))
```

@@

```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)"}
plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2")
points(
  readd(buffer_asv_pcoa),
  select = with(
    as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
    buffer == "Xpedition" & year == "2015" 
  ),
  col = "green",
  pch = 1
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "Xpedition" & year == "2016"
  ),
  col = "blue",
  pch = 2
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "LifeGuard" & year == "2016"
  ),
  col = "red",
  pch = 3
)
legend(
  "bottomright",
  legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"),
  col = c("green", "blue", "red"),
  pch = 1:3
)
# legend(
#   "topright",
#   legend = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)) %>%
#     unique() %>%
#     sort(),
#   lty = 1:3
# )
# vegan::ordiellipse(
#   readd(buffer_asv_pcoa),
#   groups = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)),
#   lty = 1:3
# )
```

@@

```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)"}
oldtheme <- theme_update(panel.border = element_blank())
readd(buffer_compare_long)
theme_set(oldtheme)
```

@@

```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)"}
readd(buffer_compare_short)
theme_set(oldtheme)

@@

```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)"}
readd(phylotax_fig)
```

@@

```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)"}
ggpubr::ggarrange(readd(conc_plot), readd(reads_plot), ncol = 2, common.legend = TRUE) +
  theme(panel.spacing = unit(0, "mm"))
```

@@

```{r qual-check, fig.cap = "(ref:qual-check-s), (ref:qual-check)", fig.scap = "(ref:qual-check-s)",fig.height=7}
ggpubr::ggarrange(
  readd(erate_plot),
  readd(eexp_plot),
  readd(eexp_region_plot),
  ncol = 1,
  # common.legend = TRUE,
  # legend = "bottom",
  labels = "auto"
)
```

@@

```{r full-length, fig.cap = "(ref:full-length-s). (ref:full-length)", fig.scap = "(ref:full-length-s)", fig.height = 6}
ggpubr::ggarrange(
  readd(full_dens_short),
  readd(full_ecdf_short),
  readd(full_dens_long),
  readd(full_ecdf_long),
  ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
)
```

@@

```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6}
readd(region_length_fig)
```

@@

```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8}
ggpubr::ggarrange(
  read_comparison(
    readd(multi_table),
    readd(comparisons),
    type = "ASV"
  ),
  read_comparison(
    readd(multi_table),
    readd(comparisons),
    type = "OTU"
  ),
  ncol = 1,
  labels = "auto",
  label.x = 0.9
)
```

@@

```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8}
theme <- theme_update(panel.border = element_blank())
readd(heattree)
theme_set(theme)
```

@@

```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7}
  readd(taxon_reads) %>%
  filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
  mutate_at("reference", replace_na, "All") %>%
  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
ggpubr::ggarrange(
  readd(tech_class_pcoa_plot),
  readd(tech_ecm_fam_pcoa_plot),
  common.legend = TRUE,
  legend = "bottom",
  labels = "auto"
)
```

@@

```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)"}
light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "method", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  filter(startsWith(method, "PHYLOTAX")) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
    amplicon = factor(amplicon, levels = c("Short", "Long"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  ggnomics::facet_nested(amplicon + tech + metric ~ guild, scales = "fixed", nest_line = TRUE) +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
  theme(strip.background = element_blank())
```

@@

```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6}
ggpubr::ggarrange(
  readd(its2_dens),
  readd(its2_ecdf),
  ncol = 1,
  labels = "auto",
  heights = c(4, 2)
)
```

@@

```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s). (ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%"}
oldtheme <- theme_update(panel.border = element_blank())
readd(heattree_length_compare)
theme_set(oldtheme)
```

@@

```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7}
readd(big_seq_table_ITS2) %>%
  colSums() %>%
  enframe(name = "ITS2", value = "nreads") %>%
  mutate(length = nchar(ITS2)) %>%
  filter(length <= 140) %>%
  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
  group_by(tech, amplicon, Algorithm, reference) %>%
  mutate(reads = reads/sum(reads)) %>%
  ungroup() %>%
  # group_by(ITS2) %>%
  # filter(!"Short" %in% amplicon) %>%
  taxon_plot(kingdom, y = reads, datasets = datasets)
```

@@

```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%"}
oldtheme <- theme_update(panel.border = element_blank())
readd(ecm_heattree)
theme_set(oldtheme)
```

@@

```{r bioinfo}
options(knitr.kable.NA = '–')

@@

```{r venn-table-asv}
readd(venn_ASV) %>%
venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)")
```

@@

```{r venn-table-otu}
readd(venn_OTU) %>%
venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)")
```

@@

```{r variofit, results="asis"}
 readd(variofit_table) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high) | mantel == FALSE,
    "-",
    sprintf(
      "%s (%s–%s)",
      formatC(estimate, digits = 2, format = "fg"),
      formatC(conf.low, digits = 2, format = "fg"),
      formatC(conf.high, digits = 2, format = "fg")
      )
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(
    names_from = "term",
    values_from = "value",
    id_cols = c("guild", "amplicon", "tech", "metric", "algorithm"),
    values_fill = list(value = "-")
  ) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

>>>>

datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccccccccicc")
regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))

gethash <- function(...) purrr::map_chr(list(...), drake_cache()$get_hash)
```

@@

```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)"}
plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2", cache = TRUE, extra.cache = gethash("buffer_asv_pcoa", "buffer_asv_table"))
points(
  readd(buffer_asv_pcoa),
  select = with(
    as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
    buffer == "Xpedition" & year == "2015" 
  ),
  col = "green",
  pch = 1
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "Xpedition" & year == "2016"
  ),
  col = "blue",
  pch = 2
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "LifeGuard" & year == "2016"
  ),
  col = "red",
  pch = 3
)
legend(
  "bottomright",
  legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"),
  col = c("green", "blue", "red"),
  pch = 1:3
)
# legend(
#   "topright",
#   legend = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)) %>%
#     unique() %>%
#     sort(),
#   lty = 1:3
# )
# vegan::ordiellipse(
#   readd(buffer_asv_pcoa),
#   groups = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)),
#   lty = 1:3
# )
```

@@

```{r buffer-compare-long, fig.cap = "(ref:buffer-compare-long)", cache = TRUE, extra.cache = gethash("buffer_compare_long")}
oldtheme <- theme_update(panel.border = element_blank())
readd(buffer_compare_long)
theme_set(oldtheme)
```

@@

```{r buffer-compare-short, fig.cap = "(ref:buffer-compare-short)", cache = TRUE, extra.cache = gethash("buffer_compare_short"), more.cache=oldtheme}
readd(buffer_compare_short)
theme_set(oldtheme)

@@

```{r phylotax, fig.height = 2, fig.width = 3, fig.cap = "(ref:phylotax-s). (ref:phylotax)", fig.scap="(ref:phylotax-s)", cache = TRUE, extra.cache = gethash("phylotax_fig")}
readd(phylotax_fig)
```

@@

```{r sample-depth, fig.cap = "(ref:sample-depth)", fig.scap = "(ref:sample-depth-s)", cache = TRUE, extra.cache = gethash("conc_plot", "reads_plot")}
ggpubr::ggarrange(readd(conc_plot), readd(reads_plot), ncol = 2, common.legend = TRUE) +
  theme(panel.spacing = unit(0, "mm"))
```

@@

```{r qual-check, fig.cap = "(ref:qual-check-s), (ref:qual-check)", fig.scap = "(ref:qual-check-s)",fig.height=7, cache = TRUE, extra.cache = gethash("erate_plot", "eexp_plot", "eexp_region_plot")}
ggpubr::ggarrange(
  readd(erate_plot),
  readd(eexp_plot),
  readd(eexp_region_plot),
  ncol = 1,
  labels = "auto"
)
```

@@

```{r full-length, fig.cap = "(ref:full-length-s). (ref:full-length)", fig.scap = "(ref:full-length-s)", fig.height = 6, cache = TRUE, extra.cache = gethash("full_dens_short", "full_ecdf_short", "full_ecdf_short", "full_ecdf_long")}
ggpubr::ggarrange(
  readd(full_dens_short),
  readd(full_ecdf_short),
  readd(full_dens_long),
  readd(full_ecdf_long),
  ncol = 1, heights = c(3, 1.7, 1.2, 2), labels = "auto", vjust = 1
)
```

@@

```{r region-lengths, fig.cap = "(ref:region-lengths-s). (ref:region-lengths)", fig.scap = "(ref:region-lengths-s)", fig.height = 6, cache = TRUE, extra.cache = gethash("region_length_fig")}
readd(region_length_fig)
```

@@

```{r read-compare, fig.cap="(ref:read-compare-s), (ref:read-compare)", fig.scap="(ref:read-compare-s)", fig.height=8, cache=TRUE, extra.cache=gethash("multi_table","comparisons")}
ggpubr::ggarrange(
  read_comparison(
    readd(multi_table),
    readd(comparisons),
    type = "ASV"
  ),
  read_comparison(
    readd(multi_table),
    readd(comparisons),
    type = "OTU"
  ),
  ncol = 1,
  labels = "auto",
  label.x = 0.9
)
```

@@

```{r taxa, fig.cap = "(ref:taxa-s). (ref:taxa)", fig.scap="(ref:taxa-s)", out.width="100%", fig.width = 8, cache=TRUE, extra.cache=gethash("heattree")}
theme <- theme_update(panel.border = element_blank())
readd(heattree)
theme_set(theme)
```

@@

```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_reads"), more.cache = datasets}
  readd(taxon_reads) %>%
  filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
  mutate_at("reference", replace_na, "All") %>%
  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)", cache=TRUE, extra.cache=gethash("tech_class_pcoa_plot", "tech_ecm_fam_pcoa_plot")}
ggpubr::ggarrange(
  readd(tech_class_pcoa_plot),
  readd(tech_ecm_fam_pcoa_plot),
  common.legend = TRUE,
  legend = "bottom",
  labels = "auto"
)
```

@@

```{r correlog, fig.cap="(ref:correlog-cap-s). (ref:correlog-cap)", fig.scap = "(ref:correlog-cap-s)", cache=TRUE, extra.cache=gethash(keep(light_plan$target, startsWith, "correlog"))}
light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "method", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  filter(startsWith(method, "PHYLOTAX")) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "W-UNIFRAC")),
    amplicon = factor(amplicon, levels = c("Short", "Long"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  ggnomics::facet_nested(amplicon + tech + metric ~ guild, scales = "fixed", nest_line = TRUE) +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1)) +
  theme(strip.background = element_blank())
```

@@

```{r ITS2-length, fig.cap = "(ref:ITS2-length-s). (ref:ITS2-length)", fig.scap = "(ref:ITS2-length-s)", fig.height = 6, cache=TRUE, extra.cache=gethash("its2_dens", "its2_ecdf")}
ggpubr::ggarrange(
  readd(its2_dens),
  readd(its2_ecdf),
  ncol = 1,
  labels = "auto",
  heights = c(4, 2)
)
```

@@

```{r heattree-length-compare, fig.cap = "(ref:heattree-length-compare-s). (ref:heattree-length-compare)", fig.scap = "(ref:heattree-length-compare-s)", out.width = "100%", cache=TRUE, extra.cache=gethash("heattree_length_compare")}
oldtheme <- theme_update(panel.border = element_blank())
readd(heattree_length_compare)
theme_set(oldtheme)
```

@@

```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7, cache=TRUE, extra.cache=gethash("big_seq_table_ITS2", "taxon_reads"), more.cache = datasets}
readd(big_seq_table_ITS2) %>%
  colSums() %>%
  enframe(name = "ITS2", value = "nreads") %>%
  mutate(length = nchar(ITS2)) %>%
  filter(length <= 140) %>%
  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
  group_by(tech, amplicon, Algorithm, reference) %>%
  mutate(reads = reads/sum(reads)) %>%
  ungroup() %>%
  # group_by(ITS2) %>%
  # filter(!"Short" %in% amplicon) %>%
  taxon_plot(kingdom, y = reads, datasets = datasets)
```

@@

```{r ecm-heattree, fig.cap = "(ref:ecm-heattree-s). (ref:ecm-heattree)", fig.scap = "(ref:ecm-heattree-s)", out.width = "100%", , cache=TRUE, extra.cache=gethash("ecm_heattree")}
oldtheme <- theme_update(panel.border = element_blank())
readd(ecm_heattree)
theme_set(oldtheme)
```

@@

```{r bioinfo, cache=TRUE, extra.cache=gethash("bioinf_table", "bioinf_header")}
options(knitr.kable.NA = '–')

@@

```{r venn-table-asv, cache=TRUE, extra.cache=gethash("venn_ASV"), more.cache=venn_table}
readd(venn_ASV) %>%
venn_table(ASVs, caption = "(ref:venn-table-asv-s). (ref:venn-table-asv)", caption.short = "(ref:venn-table-asv-s)")
```

@@

```{r venn-table-otu, cache=TRUE, extra.cache=gethash("venn_OTU"), more.cache=venn_table}
readd(venn_OTU) %>%
venn_table(OTUs, caption = "(ref:venn-table-otu-s). (ref:venn-table-otu)", caption.short = "(ref:venn-table-otu-s)")
```

@@

```{r variofit, results="asis", cache=TRUE, extra.cache=gethash("variofit_table")}
 readd(variofit_table) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high) | mantel == FALSE,
    "-",
    sprintf(
      "%s (%s–%s)",
      formatC(estimate, digits = 2, format = "fg"),
      formatC(conf.low, digits = 2, format = "fg"),
      formatC(conf.high, digits = 2, format = "fg")
      )
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(
    names_from = "term",
    values_from = "value",
    id_cols = c("guild", "amplicon", "tech", "metric", "algorithm"),
    values_fill = list(value = "-")
  ) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```
